---
output: 
  html_document: 
    keep_md: yes
---
```{r setoptions, echo=FALSE,message = FALSE,warning=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE,reindent.spaces = 2)
```  

#Activity Monitoring Analysis  
```{r library,message = FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(lattice)
```  

###Loading and preprocessing the data  
*1.Load the data*    
```{r load}
data<-read.csv("activity.csv")
```  
*2.Transform date variable to date format*    
```{r transform}
data$date<-as.Date(data$date)
```  
###What is mean total number of steps taken per day?  
*1.Calculate total number of steps per day*    
```{r total1}
data_Q1<-data%>%group_by(date)%>%summarize(steps_total=sum(steps,na.rm=T))%>%select(date,steps_total)
```   
*2.Histogram of total number of steps*    
```{r hist1}
hist(data_Q1$steps_total,breaks=seq(0,25000,2500),ylim=c(0,20),col="blue",xlab="Total num of Steps",main="Histogram of Total num of steps per Day")
```  



*3.Report Mean and Median of total number of steps*    
```{r mean1}
mean_steps<-mean(data_Q1$steps_total,na.rm=T)
median_steps<-median(data_Q1$steps_total,na.rm=T)
```    
*mean_steps:9354.23  median_steps 10395*    

###What is the average daily activity pattern?  
*1.Calculate average steps per 5-minute interval across a day*    
```{r total2}
data_Q4<-data%>%group_by(interval)%>%summarize(steps_average=mean(steps,na.rm=T))%>%select(interval,steps_average)
```    
*2.Make a time series plot of 5-minute interval and average number of steps*     
```{r hist2}
ggplot(data_Q4,aes(interval,steps_average))+geom_line(col="blue")+xlab("5-minute Interval")+
  ylab("Average Steps")+ggtitle("Average Number of Steps")+theme(plot.title = element_text(hjust = 0.5))
```  



*3.Which 5-minute interval contains the maximum number of steps?*    
```{r max}
max_index<-which.max(data_Q4$steps_average)
max_interval<-data_Q4[max_index,]$interval
```  
*max_interval:835*    

###Imputing missing values  
*1.Calculate total number of missing values*    
```{r mis}
per_step_mis<-mean(is.na(data$steps))
```  
*per_step_mis(13.11%)is missing percentage in steps variable*      
*2.Fill the missing values by the mean of corresponding 5-minute interval *   
```{r fill}
data_miss<-data
for (i in 1:dim(data_miss)[1]){
  if (is.na(data_miss[i,]$steps)){
    data_miss[i,]$steps<-data_Q4[data_Q4$interval==data_miss[i,]$interval,]$steps_average
  }
}
```  
*3.Calcluate and Create Histogram for total number of steps after missing values are imputed*    
```{r hist3}
data_miss_hist<-data_miss%>%group_by(date)%>%summarize(steps_total=sum(steps))%>%select(date,steps_total)
hist(data_miss_hist$steps_total,breaks=seq(0,25000,2500),ylim=c(0,20),col="blue",xlab="Total num of Steps",main="Histogram of Total num of steps per Day")
```  



*4.Report Mean and Median of total number of steps*     
```{r mean2}
mean_steps_miss<-mean(data_miss_hist$steps_total,na.rm=T)
median_steps_miss<-median(data_miss_hist$steps_total,na.rm=T)
```
*mean_steps_miss:10766.19  median_steps_miss:10766.19*  

###Are there differences in activity patterns between weekdays and weekends?    
*1.Create day variable that contain two levels-"weekday" and "weekend";Calculate average steps for 5-minute interval across weekday or weekend*    
```{r weekday}
data_miss$day<-ifelse(weekdays(data_miss$date)%in%c("Monday","Tuesday","Thursday","Wednesday","Friday"),"Weekday","Weekend")
average_step<-data_miss%>%group_by(interval,day)%>%summarize(steps_mean=mean(steps))%>%select(day,steps_mean,interval)
```  
*2.Make Panel Plot*    
```{r panel}
xyplot(steps_mean~interval|day,average_step,
       type="l",
       lwd=1,
       xlab="Interval",
       ylab="Number of Steps",
       layout=c(1,2))
```